Syntax-Based Concept Alignment for Domain-Specific Machine Translation

### 2021-05-09 paper/paper.tex
finished CE eval excepts parts depending on experiments to run

% TODO: add figure comapring tree representations

@@

\section{Evaluation} \label{evaluation}
\subsection{Data}
Because we want part of our evaluation to be independent from the quality of UD parsing, some of the experiments are carried out using manually annotated data. 
To this purpose, we use a 100-sentences subset of the Parallel UD (PUD) corpus, a set of handcrafted\footnote{Annotation is sometimes semi-automatic, but all treebanks are at least manually checked.} multilingual treebanks in CoNLL-U format created for the CoNLL 2017 shared task on Multilingual Parsing from Raw Text to Universal Dependencies \cite{TODO:}.
PUD treebanks are available in over 20 languages, of which we selected English, Italian and Swedish. 
Using this data limits the amount of alignment errors that are due annotation issues to a minimum, even though a small number of inconsistencies and imprecisions is present even in this corpus.

@@

@@

\subsubsection{Results} \label{ceres}
% PUD (incl. reason-wise?): independent, comparison with SMT
% RAW
\subsubsection{Discussion}

\subsection{MT experiments}
% metrics & methods
\subsubsection{Results}
\subsubsection{Discussion}

>>>>

% TODO: add figure comparing tree representations

@@

\section{Evaluation} \label{evaluation}
% TODO: move around things to have something here (or outline structure of the section)
\subsection{Data}
Because we want part of our evaluation to be independent from the quality of UD parsing, some of the experiments are carried out using manually annotated data. 
To this purpose, we use a 100-sentences subset of the Parallel UD (PUD) corpus, a set of handcrafted\footnote{Annotation is sometimes semi-automatic, but all treebanks are at least manually checked.} multilingual treebanks in CoNLL-U format created for the CoNLL 2017 shared task on Multilingual Parsing from Raw Text to Universal Dependencies \cite{TODO:}.
PUD treebanks are available in over 20 languages, of which we selected English, Italian and Swedish. 
Using this data limits the amount of alignment errors that are due annotation issues to a minimum, even though a small number of inconsistencies and imprecisions is present even in this corpus.

@@

% TODO: size

@@

\subsubsection{Results on manually annotated treebanks} \label{ceres}
In table \ref{pud_fast}, we compare the results obtained with our grammar-based module to those obtained statistically on the PUD treebanks.

Of course, \texttt{fast\_align} does not make use of the information present in the CoNLL-U files excepts with regards to tokenization. 
On the other hand, the relatively large size of the PUD treebanks makes it possible to also train the statistical tool on the full dataset instead of just using the chosen 100-sentences subset, allowing for a fairer comparison.

In both cases, \texttt{fast\_align} is run with 10 iterations in EM training, symmetrizing the alignments found by running the program in both directions.

%TODO: position properly
\begin{table*}[h]
  \centering
  \small
  \begin{tabular}{l|lll}
  %\hline
   & \textbf{CE} & \textbf{\texttt{fast\_align} 100} & \textbf{\texttt{fast\_align} 1000}\\ \hline
   \textbf{tot. alignments} & 716 & 1440 & 1435 \\ 
   \textbf{\texttt{+} and \texttt{=}} & 536 (75\%) & 410 (28\%) & 656 (46\%)\\ 
   \textbf{\texttt{+}} & 491 (69\%) & 371 (26\%) & 590 (41\%)\\ 
  \end{tabular}
  \caption[Comparison between our grammar-based CE module and \texttt{fast\_align}]{Comparison between our grammar-based CE module and \texttt{fast\_align} on PUD data. The middle column shows the results obtained by training \texttt{fast\_align} on the 100-sentence subset fed to the extraction module. On the right, the results obtained by training \texttt{fast\_align} on the full PUD treebanks dataset, discarding the alignments obtained for sentences 101-1000 for the final comparison. Row 1 reports the total number of distinct alignments obtain. Rows 2 and 3 give the percentages of correct alignments, with and without considering sentence pair-specific correspondences as correct.}
  \label{pud_fast}
\end{table*}

While \texttt{fast\_align} is designed to align every word in the text (or explicitly state that a word has no counterpart in its translation) and, consequently, extracts around twice as many correspondences as our CE module, the percentage of correct correspondences is definitely in favor of our system, even though \texttt{fast\_align} gets significantly more precise when trained on the full, 1000-sentence dataset.

\subsubsection{Results on raw text}

The course plan corpora are significantly harder to work with, both for our CE module and \texttt{fast\_align}.
In the former case, the additional challenge is relying on the quality of automatic parsing, while \texttt{fast\_align} suffers from the small size of the datasets, which are fed in their entirety to both systems.

The results of this comparison are summarized in table \ref{raw_fast}.

%TODO: position properly
%TODO: run experiments & update data
\begin{table*}[h]
  \centering
  \small
  \begin{tabular}{l|lll}

  \end{tabular}
  \caption[Comparison between our grammar-based CE module and \texttt{fast\_align}]{Comparison between our grammar-based CE module and \texttt{fast\_align} on raw data, using the DMI and CSE corpora. Row 1 reports the total number of distinct alignments obtain. Rows 2 and 3 give the percentages of correct alignments, with and without considering sentence pair-specific correspondences as correct.}
  \label{raw_fast}
 \end{table*}

 % TODO: discussion

% PUD: incl. reason-wise?)

\subsection{MT experiments}
% metrics & methods
\subsubsection{Experimental results}
